---
title: "Penguins size comparisons"
author: "Roshni Katrak-Adefowora"
date: "11/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

#Part 1. T-Tests

Mass comparison: male & female chinstrap penguins
```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

#Histograms
ggplot(data=chinstraps, aes(x=body_mass_g))+
  geom_histogram(bins=10)+
  facet_wrap(~sex)

#QQ plot - looks very linear - very normal!
ggplot(data=chinstraps, aes(sample=body_mass_g))+
  geom_qq()+
  facet_wrap(~sex)

#Get descriptive statistics (summary table)
chinstraps %>% 
  group_by(sex) %>% 
  summarize(mean_mass = mean(body_mass_g),
            median_mass = median(body_mass_g),
            sd_mass = sd(body_mass_g),
            sample_size = n())
```

Get vectors of body_mass_g for female and male chinstrap penguins separately
```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

### T-tests, a couple different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins? (two-sided t-test)
```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

#effect size (get 1.26, large difference)
cohen.d(chinstraps_f, chinstraps_m)
```

2. Are male chinstrap penguins significantly larger in body mass than female chinstrap penguins? (one-sided t-test)
```{r}
t.test(chinstraps_m, chinstraps_f, alternative="greater") #use "greater" or "less," but should line up with order the vectors are in
```

