---
title: "Penguins size comparisons"
author: "Roshni Katrak-Adefowora"
date: "11/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

#Part 1. T-Tests

Mass comparison: male & female chinstrap penguins
```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

#Histograms
ggplot(data=chinstraps, aes(x=body_mass_g))+
  geom_histogram(bins=10)+
  facet_wrap(~sex)

#QQ plot - looks very linear - very normal!
ggplot(data=chinstraps, aes(sample=body_mass_g))+
  geom_qq()+
  facet_wrap(~sex)

#Get descriptive statistics (summary table)
chinstraps %>% 
  group_by(sex) %>% 
  summarize(mean_mass = mean(body_mass_g),
            median_mass = median(body_mass_g),
            sd_mass = sd(body_mass_g),
            sample_size = n())
```

Get vectors of body_mass_g for female and male chinstrap penguins separately
```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

### T-tests, a couple different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins? (two-sided t-test)
```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

#effect size (get 1.26, large difference)
cohen.d(chinstraps_f, chinstraps_m)
```

2. Are male chinstrap penguins significantly larger in body mass than female chinstrap penguins? (one-sided t-test)
```{r}
t.test(chinstraps_m, chinstraps_f, alternative="greater") #use "greater" or "less," but should line up with order the vectors are in
```

## One-way ANOVA

## 1. Table of counts by species and sex

```{r}
penguin_counts <- penguins %>% 
  count(species, sex)

penguin_counts
```

Create a subset of only female penguins
```{r}
penguin_f <- penguins %>% 
  filter(sex == "female")

#Histogram
ggplot(data=penguin_f, aes(x=flipper_length_mm))+
  geom_histogram(aes(fill=species))+
  facet_wrap(~species, scales="free") #scales command makes it so that they're not all on the same scale
#look pretty normal

#QQ plot
ggplot(data=penguin_f, aes(sample=flipper_length_mm))+
  geom_qq(aes(color=species))+
  facet_wrap(~species, scales="free")
```

### Summary statistics

Starting from penguin_f, create a summary table that groups by species, and returns the following:

- mean flipper length
- sd
- sample size
- variance `var`
- standard error = sd/sqrt(n)

```{r}
penguin_f_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarize(mean_flipper_length=mean(flipper_length_mm),
            sd_flipper_length=sd(flipper_length_mm),
            sample_size=n(),
            se_flipper_length=sd(flipper_length_mm)/sqrt(n()),
            var_flipper_length=var(flipper_length_mm))

leveneTest(flipper_length_mm ~ species, data=penguin_f) #variance in flipper length as function of species, data exists in penguin_f
#p>0.05, retain null hypothesis that variances are equal
```

### Beeswarm plot with mean +- 1sd

```{r}
ggplot()+
  geom_beeswarm(data=penguin_f, 
                aes(x=species, y=flipper_length_mm),
                color="purple",
                size=3,
                alpha=0.5)+ #transparency
  geom_point(data=penguin_f_summary, #add mean points
             aes(x=species, y=mean_flipper_length),
             size=4)+
  geom_errorbar(data=penguin_f_summary, #add whiskers, mean +- 1sd
                aes(x=species,
                    ymax=mean_flipper_length + sd_flipper_length,
                ymin=mean_flipper_length - sd_flipper_length),
                width=0.1)+
  theme_light()
```

### One-way ANOVA

```{r}
#aov function by default runs type 1 ANOVA, but you want type 3 if very unequal sample sizes/groups/distributions (usually they give similar results, unless you have very unbalanced data/design)
penguin_aov <- aov(flipper_length_mm ~ species, data=penguin_f) 
summary(penguin_aov)

#post-hoc test, does pairwise comparisons
TukeyHSD(penguin_aov)
```

Type III ANOVA with the car package
```{r}
#create linear model
penguin_lm <- lm(flipper_length_mm ~ species, data=penguin_f)
penguin_aov_t3 <- Anova(penguin_lm, type="III")
penguin_aov_t3
```

